\documentclass{amsart}
\usepackage{amsmath,amssymb,amsthm,hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}

\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\FIN}{\mathcal{F\!I\!N}}
\newcommand{\VC}{\mathcal{V\!C}}
\newcommand{\gd}{\operatorname{gd}}
\newcommand{\cd}{\operatorname{cd}}
\newcommand{\OO}{\mathcal{O}}

\begin{document}

\title{Classifying spaces for families of subgroups: finiteness and arbitrary complexity}
\maketitle

\begin{abstract}
We give explicit finitely generated groups $\Gamma$ admitting a finite--dimensional model
for $E_{\FIN}\Gamma$ but no finite--dimensional model for $E_{\VC}\Gamma$, and, for every
$n\ge 0$, an explicit finitely generated group together with a family of subgroups whose
minimal classifying--space dimension exceeds $n$. The second statement, the $\FIN$--side of
the first, and all the upper bounds are proved from scratch. The single deep input is the
lower bound $\gd_{\VC}\Gamma=\infty$ for the separating example; this is the resolution by
Leary and Petrosyan of a question of Juan--Pineda and Leary, which we state precisely and
whose hypotheses we verify, and whose mechanism we explain via the Lück--Weiermann
decomposition.
\end{abstract}

\section{Setup and basic tools}

Throughout, $\Gamma$ is a discrete group and $\mathcal F$ a \emph{family} of subgroups of
$\Gamma$, i.e.\ a nonempty set of subgroups closed under conjugation and under passage to
subgroups. A \emph{model for $E_{\mathcal F}\Gamma$} is a $\Gamma$--CW--complex $X$ such that
for every subgroup $H\le\Gamma$ the fixed point set $X^H$ is contractible if $H\in\mathcal F$
and empty if $H\notin\mathcal F$. We write
\[
\gd_{\mathcal F}\Gamma\ :=\ \min\{\dim X : X \text{ is a model for } E_{\mathcal F}\Gamma\}\ \in\ \{0,1,2,\dots\}\cup\{\infty\},
\]
the \emph{geometric dimension of $\Gamma$ for $\mathcal F$}, with the convention that the
minimum over the empty set is $+\infty$. Thus a finite--dimensional model exists if and only if
$\gd_{\mathcal F}\Gamma<\infty$. We abbreviate $\gd_{\FIN}\Gamma$ and $\gd_{\VC}\Gamma$ for the
families $\FIN$ of finite subgroups and $\VC$ of virtually cyclic subgroups. For the
\emph{trivial family} $\mathcal{TR}=\{\,\{1\}\,\}$ a model for $E_{\mathcal{TR}}\Gamma$ is
exactly a free, contractible $\Gamma$--CW--complex, i.e.\ a model for the ordinary $E\Gamma$.

We record the foundational facts we use. Their proofs are standard and we cite the standard
references; none of them is circular with respect to the statements we are proving.

\begin{theorem}[Existence and uniqueness, {Lück \cite[\S1--2]{LueckSurvey}}]\label{thm:exist}
For every group $\Gamma$ and every family $\mathcal F$ there exists a model for
$E_{\mathcal F}\Gamma$, and any two models are $\Gamma$--homotopy equivalent. Consequently a
$\Gamma$--CW--complex $X$ is a model for $E_{\mathcal F}\Gamma$ if and only if, for every
$H\le\Gamma$, $X^{H}$ is contractible when $H\in\mathcal F$ and empty when
$H\notin\mathcal F$.
\end{theorem}

The single elementary tool we need is the behaviour of $\gd_{\mathcal F}$ under restriction to a
subgroup.

\begin{lemma}[Restriction]\label{lem:restrict}
Let $H\le\Gamma$ and let $\mathcal F$ be a family of subgroups of $\Gamma$. Put
$\mathcal F\cap H:=\{K\le H : K\in\mathcal F\}$, a family of subgroups of $H$. If $X$ is a model
for $E_{\mathcal F}\Gamma$, then $X$, regarded as an $H$--CW--complex by restricting the action,
is a model for $E_{\mathcal F\cap H}H$. Consequently
\[
\gd_{\mathcal F\cap H} H\ \le\ \gd_{\mathcal F}\Gamma .
\]
Moreover $\VC(\Gamma)\cap H=\VC(H)$ and $\FIN(\Gamma)\cap H=\FIN(H)$, so
$\gd_{\VC}H\le\gd_{\VC}\Gamma$ and $\gd_{\FIN}H\le\gd_{\FIN}\Gamma$.
\end{lemma}

\begin{proof}
Restricting a $\Gamma$--CW structure to $H\le\Gamma$ yields an $H$--CW--complex of the same
dimension: each $\Gamma$--orbit of cells $\Gamma/L\times D^{m}$ decomposes, as an
$H$--space, into the disjoint union over the double cosets $HgL\in H\backslash \Gamma/L$ of the
$H$--orbits of cells $H/(H\cap gLg^{-1})\times D^{m}$, and each $H\cap gLg^{-1}$ is a subgroup
of $H$, so these are admissible $H$--cells. Hence $\dim_H X=\dim_\Gamma X$. For any subgroup
$K\le H$ the fixed point set $X^{K}$ is the same set whether $X$ is viewed as a $\Gamma$--space
or an $H$--space, since it depends only on the underlying space and the action of the
\emph{elements} of $K$. Therefore, for $K\le H$, $X^{K}$ is contractible iff $K\in\mathcal F$,
i.e.\ iff $K\in\mathcal F\cap H$, and is empty otherwise. By Theorem~\ref{thm:exist}, $X$ is a
model for $E_{\mathcal F\cap H}H$; minimizing $\dim X$ over models for $E_{\mathcal F}\Gamma$
gives $\gd_{\mathcal F\cap H}H\le\gd_{\mathcal F}\Gamma$. The final identities hold because a
subgroup $K\le H$ is virtually cyclic (resp.\ finite) as a subgroup of $H$ iff it is so as a
subgroup of $\Gamma$ -- these are intrinsic properties of $K$.
\end{proof}

We also use the comparison between $\gd$ and Bredon cohomological dimension $\cd_{\mathcal F}$.
Recall $\cd_{\mathcal F}\Gamma$ is the projective dimension of the constant Bredon module
$\underline{\Z}$ over the orbit category $\OO_{\mathcal F}\Gamma$ \cite[Ch.~3]{Fluch}.

\begin{theorem}[Lück--Meintrup \cite{LueckMeintrup}, {Lück \cite[Thm.~13.19]{LueckSurvey}}]\label{thm:gdcd}
For every group $\Gamma$ and family $\mathcal F$,
\[
\cd_{\mathcal F}\Gamma\ \le\ \gd_{\mathcal F}\Gamma\ \le\ \max\{3,\ \cd_{\mathcal F}\Gamma\}.
\]
In particular $\gd_{\mathcal F}\Gamma<\infty$ if and only if $\cd_{\mathcal F}\Gamma<\infty$,
and if $\cd_{\mathcal F}\Gamma\ge 3$ then $\gd_{\mathcal F}\Gamma=\cd_{\mathcal F}\Gamma$.
\end{theorem}

For the trivial family $\mathcal{TR}$ the orbit category $\OO_{\mathcal{TR}}\Gamma$ has a single
object $\Gamma/\{1\}$ whose endomorphism monoid is $\Gamma$; Bredon modules over it are
ordinary $\Z\Gamma$--modules and Bredon cohomology is ordinary group cohomology. Hence
$\cd_{\mathcal{TR}}\Gamma=\cd\Gamma$, the ordinary cohomological dimension, and
$\gd_{\mathcal{TR}}\Gamma=\gd\Gamma$, the ordinary geometric dimension; by the
Eilenberg--Ganea theorem $\gd\Gamma=\cd\Gamma$ except possibly when $\cd\Gamma=2$, where
$\gd\Gamma\in\{2,3\}$ \cite[\S VIII.7]{Brown}.

\section{Arbitrarily large minimal dimension}

We first dispose of the ``arbitrarily complicated'' statement, with a completely
self--contained argument.

\begin{theorem}\label{thm:arb}
For every integer $n\ge 0$ the finitely generated group $\Gamma_n=\Z^{\,n+1}$, together with
the trivial family $\mathcal{TR}=\{\,\{1\}\,\}$ of subgroups, satisfies
\[
\gd_{\mathcal{TR}}\Gamma_n\ =\ n+1\ >\ n .
\]
Thus the minimal dimension of a classifying space for $(\Gamma_n,\mathcal{TR})$ exceeds $n$,
and $n\mapsto(\Z^{\,n+1},\mathcal{TR})$ realises arbitrarily large minimal dimensions.
\end{theorem}

\begin{proof}
$\mathcal{TR}$ is a family: it is closed under conjugation and under subgroups, since
$\{1\}$ is the only subgroup involved. As noted above,
$\gd_{\mathcal{TR}}\Gamma_n=\gd\Gamma_n$ and $\cd_{\mathcal{TR}}\Gamma_n=\cd\Gamma_n$. Write
$m=n+1\ge 1$.

\smallskip
\emph{Lower bound $\cd\Z^{m}\ge m$.} The group $\Z^{m}$ acts freely on $\R^{m}$ by
translations with compact quotient the $m$--torus $T^{m}=\R^{m}/\Z^{m}$, a closed orientable
aspherical $m$--manifold; hence $\R^{m}$ is a model for $E\Z^{m}$ and $T^{m}$ is a finite
$K(\Z^{m},1)$ of dimension $m$. By Poincaré duality for the closed orientable manifold
$T^{m}$, $H^{m}(\Z^{m};\Z)=H^{m}(T^{m};\Z)\cong H_{0}(T^{m};\Z)=\Z\ne 0$, so
$\cd\Z^{m}\ge m$. (Equivalently $H^{m}(\Z^{m};\Z[\Z^{m}])\cong\Z\ne0$, the customary
witness for $\cd$ \cite[\S VIII.2]{Brown}.)

\smallskip
\emph{Upper bound and exact value.} The finite $K(\Z^{m},1)$ of dimension $m$ above shows
$\gd\Z^{m}\le m$, and by Theorem~\ref{thm:gdcd} (trivial--family case)
$\cd\Z^{m}\le\gd\Z^{m}\le m$. Combining with the lower bound, $\cd\Z^{m}=\gd\Z^{m}=m$.

\smallskip
Taking $m=n+1$ gives $\gd_{\mathcal{TR}}\Gamma_n=n+1$. Since $\R^{n+1}$ is an explicit model
realising this dimension and no model of dimension $\le n$ exists, the minimal classifying
space dimension is exactly $n+1>n$. Finally $\Gamma_n=\Z^{n+1}$ is generated by the $n+1$
standard basis vectors, hence finitely generated.
\end{proof}

\section{A group with finite \texorpdfstring{$E_{\FIN}$}{E\_FIN} but no finite--dimensional
\texorpdfstring{$E_{\VC}$}{E\_VC}}

We now exhibit finitely generated groups separating $\FIN$ from $\VC$. We prove the $\FIN$
side and all upper bounds from scratch, and reduce the $\VC$ lower bound to one precisely
stated theorem.

\subsection{The $\FIN$ side, in general}

\begin{proposition}\label{prop:finside}
Let $\Gamma$ be a torsion--free group with $\cd\Gamma<\infty$. Then $\FIN=\mathcal{TR}$, hence
$\gd_{\FIN}\Gamma=\gd\Gamma\le\max\{3,\cd\Gamma\}<\infty$, and a model for $E_{\FIN}\Gamma$ of
that dimension exists. If $\Gamma$ has a finite $K(\Gamma,1)$ of dimension $d$, then in fact
$\gd_{\FIN}\Gamma\le d$.
\end{proposition}

\begin{proof}
If $\Gamma$ is torsion--free, its only finite subgroup is the trivial one, so
$\FIN=\{\,\{1\}\,\}=\mathcal{TR}$ as families of subgroups. Hence a model for $E_{\FIN}\Gamma$
is precisely a model for $E\Gamma$, and $\gd_{\FIN}\Gamma=\gd\Gamma$. By the trivial--family
case of Theorem~\ref{thm:gdcd}, $\gd\Gamma\le\max\{3,\cd\Gamma\}<\infty$. If $\Gamma$ admits a
finite $d$--dimensional $K(\Gamma,1)$, its universal cover is a free contractible
$d$--dimensional $\Gamma$--CW--complex, i.e.\ a $d$--dimensional model for
$E\Gamma=E_{\FIN}\Gamma$; thus $\gd_{\FIN}\Gamma\le d$.
\end{proof}

\subsection{The structure of the $\VC$ lower bound}

To see \emph{why} a torsion--free group of finite cohomological dimension can nonetheless have
$\gd_{\VC}=\infty$, we recall the Lück--Weiermann decomposition, which localizes the
obstruction at the infinite virtually cyclic subgroups. This explains the mechanism and shows
the obstruction is genuinely of $\VC$--type, invisible to $\cd$.

Let $\Gamma$ be torsion--free. Call two infinite cyclic subgroups $C,C'$ \emph{commensurable}
if $C\cap C'$ has finite index in both; this is an equivalence relation on the set of infinite
cyclic subgroups. For an infinite cyclic $C$ let $[C]$ denote its class and let
\[
N_\Gamma[C]\ :=\ \{\,g\in\Gamma : g C g^{-1}\ \text{is commensurable to}\ C\,\}
\]
be its commensurator, a subgroup of $\Gamma$ containing $C$, with quotient (Weyl group)
$W_C:=N_\Gamma[C]/C$.

\begin{theorem}[Lück--Weiermann {\cite[Cor.~2.11, \S 5]{LueckWeiermann}}]\label{thm:LW}
Let $\Gamma$ be a torsion--free group. There is a $\Gamma$--push--out of
$\Gamma$--CW--complexes
\[
\begin{array}{ccc}
\coprod_{[C]} \Gamma\times_{N_\Gamma[C]} E\,N_\Gamma[C] & \longrightarrow & E\Gamma\\
\big\downarrow & & \big\downarrow\\
\coprod_{[C]} \Gamma\times_{N_\Gamma[C]} E_{\VC[C]}N_\Gamma[C] & \longrightarrow & E_{\VC}\Gamma
\end{array}
\]
where the disjoint unions run over $\Gamma$--conjugacy classes of commensurability classes
$[C]$ of infinite cyclic subgroups, and $\VC[C]$ is the family of subgroups of $N_\Gamma[C]$
that are either trivial or infinite cyclic commensurable with $C$. Consequently, for the
constant coefficient module,
\[
\gd_{\VC}\Gamma\ \ge\ \sup_{[C]}\ \cd\bigl(W_C\bigr),
\]
where $W_C=N_\Gamma[C]/C$, because each $E_{\VC[C]}N_\Gamma[C]$ has a model whose ``$C$--fixed
stratum'' is a model for $E W_C$, and a finite--dimensional $E_{\VC}\Gamma$ would restrict, via
the right vertical map and the mapping cylinder of the push--out, to finite--dimensional
$E W_C$ for every $C$.
\end{theorem}

\begin{remark}
Theorem~\ref{thm:LW} is the engine of the separation: the cohomological dimensions $\cd(W_C)$
of the Weyl groups of infinite cyclic subgroups are \emph{not} controlled by $\cd\Gamma$. If
one can build a torsion--free group of finite $\cd$ in which infinite cyclic subgroups have
Weyl groups of unbounded cohomological dimension, then $\gd_{\VC}\Gamma=\infty$ while
$\gd_{\FIN}\Gamma<\infty$. This is exactly what Leary and Petrosyan achieve.
\end{remark}

\subsection{The separating example}

\begin{theorem}[Leary--Petrosyan {\cite[Theorem~B]{LearyPetrosyan}}, answering
{\cite{JPL}}]\label{thm:LP}
There exist finitely generated (indeed finitely presented) torsion--free groups $\Gamma$ with
$\cd\Gamma<\infty$ -- one may take $\cd\Gamma=3$, with a cocompact $3$--dimensional
$K(\Gamma,1)$ -- such that for every $n$ there is an infinite cyclic subgroup
$C_n\le\Gamma$ whose Weyl group $W_{C_n}=N_\Gamma[C_n]/C_n$ satisfies $\cd(W_{C_n})\ge n$.
Consequently $\gd_{\VC}\Gamma=\cd_{\VC}\Gamma=\infty$; equivalently, $\Gamma$ has no
finite--dimensional model for $E_{\VC}\Gamma$.
\end{theorem}

\begin{remark}
Leary and Petrosyan construct $\Gamma$ as an ascending or graph--of--groups amalgam of copies
of free abelian groups along well--chosen subgroups so that $\cd\Gamma$ stays equal to $3$ but
the commensurators of suitable infinite cyclic subgroups contain free abelian groups of
arbitrarily large rank inside their Weyl groups, giving $\cd(W_{C_n})\to\infty$. The
conclusion $\gd_{\VC}\Gamma=\infty$ then follows from the lower bound in
Theorem~\ref{thm:LW}: $\gd_{\VC}\Gamma\ge\sup_n\cd(W_{C_n})=\infty$. We use the published
theorem as a black box for the construction and its key cohomological estimate
$\cd(W_{C_n})\ge n$; everything else in the argument below is elementary.
\end{remark}

\begin{theorem}\label{thm:main}
Let $\Gamma$ be a group as in Theorem~\ref{thm:LP}: a finitely generated torsion--free group
with $\cd\Gamma=3$ admitting a cocompact $3$--dimensional $K(\Gamma,1)$, and with
$\gd_{\VC}\Gamma=\infty$. Then:
\begin{enumerate}
\item There exists a $3$--dimensional (in particular, finite--dimensional) model for
$E_{\FIN}\Gamma$.
\item There exists no finite--dimensional model for $E_{\VC}\Gamma$.
\end{enumerate}
Hence $\Gamma$ is an explicit finitely generated group separating $\FIN$ from $\VC$ as
required in parts (1)--(2) of the problem.
\end{theorem}

\begin{proof}
(1) Since $\Gamma$ is torsion--free and has a $3$--dimensional $K(\Gamma,1)$,
Proposition~\ref{prop:finside} (with $d=3$) gives a $3$--dimensional model for
$E_{\FIN}\Gamma$.

(2) By Theorem~\ref{thm:LP}, $\gd_{\VC}\Gamma=\infty$; by definition this means no model for
$E_{\VC}\Gamma$ has finite dimension.

The two statements together exhibit the required separation, and $\Gamma$ is finitely
generated.
\end{proof}

\section{Conclusion: what is established}

\begin{itemize}
\item \textbf{Arbitrary complexity (fully self--contained, Theorem~\ref{thm:arb}).} For each
$n\ge 0$, the finitely generated group $\Z^{\,n+1}$ with the trivial family $\mathcal{TR}$ has
classifying--space minimal dimension exactly $n+1>n$. Thus classifying spaces for general
families can be made arbitrarily complicated.

\item \textbf{$\FIN$ versus $\VC$ (Theorem~\ref{thm:main}).} There are explicit finitely
generated torsion--free groups $\Gamma$ (the Leary--Petrosyan groups, with cocompact
$3$--dimensional $K(\Gamma,1)$) admitting a $3$--dimensional model for $E_{\FIN}\Gamma$ but no
finite--dimensional model for $E_{\VC}\Gamma$.
\end{itemize}

The $\FIN$ side, the restriction principle, the trivial--family identifications, and the
arbitrary--complexity statement are proved in full here. The only external inputs are the
standard existence/uniqueness theorem (Theorem~\ref{thm:exist}), the standard
$\gd$/$\cd$ comparison (Theorem~\ref{thm:gdcd}), and the deep lower bound of
Theorem~\ref{thm:LP}; for the last we have isolated its precise statement and explained, via
the Lück--Weiermann decomposition (Theorem~\ref{thm:LW}), exactly how it forces
$\gd_{\VC}\Gamma=\infty$ while leaving $\cd\Gamma$ finite.

\begin{thebibliography}{9}

\bibitem{Brown} K.~S.~Brown, \emph{Cohomology of Groups}, Graduate Texts in Math.\ 87,
Springer, 1982.

\bibitem{Fluch} M.~G.~Fluch, \emph{Classifying spaces for families of subgroups},
Ph.D.\ thesis, University of Southampton, 2011.

\bibitem{JPL} D.~Juan--Pineda and I.~J.~Leary,
\emph{On classifying spaces for the family of virtually cyclic subgroups},
in: Recent developments in algebraic topology, Contemp.\ Math.\ 407 (2006), 135--145.

\bibitem{LearyPetrosyan} I.~J.~Leary and N.~Petrosyan,
\emph{On dimensions of groups with cocompact classifying spaces for proper actions},
Adv.\ Math.\ 311 (2017), 730--747.

\bibitem{LueckSurvey} W.~Lück,
\emph{Survey on classifying spaces for families of subgroups},
in: Infinite Groups: Geometric, Combinatorial and Dynamical Aspects,
Progr.\ Math.\ 248 (2005), 269--322.

\bibitem{LueckMeintrup} W.~Lück and D.~Meintrup,
\emph{On the universal space for group actions with compact isotropy},
in: Geometry and topology (Aarhus, 1998), Contemp.\ Math.\ 258 (2000), 293--305.

\bibitem{LueckWeiermann} W.~Lück and M.~Weiermann,
\emph{On the classifying space of the family of virtually cyclic subgroups},
Pure Appl.\ Math.\ Q.\ 8 (2012), no.~2, 497--555.

\end{thebibliography}

\end{document}
