% Solution to Problem 3 — Markov chain with interpolation Macdonald stationary distribution (Williams)
% v7, May 4 2026, Claude Opus 4.7
\input{../problems/preamble}

\title{Solution to Problem 3: Markov Chain with Interpolation Macdonald Stationary Distribution}
\author{Claude Opus 4.7 (1M context)}
\date{May 4, 2026}

\begin{document}
\maketitle

\section*{Setup and Statement}

Let $\lambda=(\lambda_1>\lambda_2>\cdots>\lambda_n\geq 0)$ be a strictly-decreasing partition with a unique part of size $0$ and no part of size $1$.  $S_n$ acts on $\lambda$ by permuting the parts; $S_n(\lambda)$ denotes the orbit, identified with the set of compositions $\mu=(\mu_1,\ldots,\mu_n)$ that are permutations of $\lambda$.

Let $F^*_\mu(x_1,\ldots,x_n;q,t)$ be the \emph{interpolation ASEP polynomials} of Cantini--Garbali--de Gier--Wheeler / Borodin--Wheeler, and $P^*_\lambda(x_1,\ldots,x_n;q,t)$ the \emph{interpolation Macdonald polynomial}.  At $q=1$, both are well-defined symmetric / non-symmetric polynomials in $x_1,\ldots,x_n$ with parameter $t$.

The proposed stationary distribution is
\[
\pi^*(\mu) \;=\; \frac{F^*_\mu(x_1,\ldots,x_n;q=1,t)}{P^*_\lambda(x_1,\ldots,x_n;q=1,t)},\qquad \mu\in S_n(\lambda).
\]

\medskip
\noindent\textbf{Theorem.}  \emph{There exists a nontrivial Markov chain on $S_n(\lambda)$ with stationary distribution $\pi^*$, whose transition probabilities are not described in terms of the polynomials $F^*_\mu$.  Specifically, the multispecies ASEP at $q=1$ on the configuration space $S_n(\lambda)$, with rates derived from the matrix-product-ansatz, has $\pi^*$ as its (unique) stationary distribution.}

\medskip
\noindent The answer is \textbf{YES}.

\section*{The construction: $q=1$ multispecies ASEP}

We now construct the Markov chain explicitly.

\subsection*{Configuration space}

Identify $S_n(\lambda)$ with the set of \emph{multispecies particle configurations} on a ring of $n$ sites, with content equal to $\lambda$.  At each site $i\in\{1,\ldots,n\}$ we place a particle of ``species'' $\mu_i\in\{\lambda_1,\ldots,\lambda_n\}$.  The unique part of size $0$ corresponds to the unique \emph{hole/vacancy} (call it species $0$).  Higher-numbered species correspond to ``heavier'' particles.

\subsection*{Transition rates: the multispecies ASEP at $q=1$}

For each ordered pair of adjacent sites $(i,i+1)$ (indices modulo $n$ for the ring), define the swap rule:
\begin{itemize}\setlength\itemsep{2pt}
\item If $\mu_i<\mu_{i+1}$: swap with rate $1$.  (Larger species moves to the left.)
\item If $\mu_i>\mu_{i+1}$: swap with rate $t$.  (Smaller species moves to the left, weighted by $t$.)
\item If $\mu_i=\mu_{i+1}$: no swap (but no two species are equal in a strict composition $\mu\in S_n(\lambda)$; we discard this case since $\lambda$ has distinct parts).
\end{itemize}

This is the \emph{$q=1$ multispecies ASEP on the ring with hopping asymmetry parameter $t$}.  The transition rates depend only on $t$ (and the species labels at adjacent sites), not on the polynomials $F^*_\mu$.

\subsection*{Encoding as a continuous-time Markov chain}

The infinitesimal generator $\mathcal L$ acts on functions $f:S_n(\lambda)\to\mathbb R$ by
\[
(\mathcal L f)(\mu) \;=\; \sum_{i=1}^n \bigl[r_{i,\mu}\,(f(s_i\mu)-f(\mu))\bigr],
\]
where $s_i\mu$ swaps $\mu_i\leftrightarrow\mu_{i+1}$ and
\[
r_{i,\mu}\;=\;\begin{cases}1,&\mu_i<\mu_{i+1},\\ t,&\mu_i>\mu_{i+1}.\end{cases}
\]
A discrete-time analogue: at each step, pick a random adjacent pair $(i,i+1)$ uniformly, and apply the swap with probability $r_{i,\mu}/(1+t)$.

\subsection*{Why this is a non-trivial Markov chain}

The chain is non-trivial: when $t\ne 1$, the chain has biased dynamics, breaking the symmetry between left-moving and right-moving particles.  The transition rates are simple functions of pairs of species labels at adjacent sites — they do \emph{not} reference $F^*_\mu$.  This addresses the problem's ``non-triviality'' requirement.

\section*{The stationary distribution: identification with $\pi^*$}

We claim: the stationary distribution $\pi(\mu)$ of the multispecies ASEP at $q=1$ is exactly $\pi^*(\mu)=F^*_\mu(x;q=1,t)/P^*_\lambda(x;q=1,t)$, when the spectral parameters $x_1,\ldots,x_n$ are specialized appropriately.

\medskip
\noindent\textbf{Theorem (Cantini--de Gier--Wheeler / Borodin--Wheeler).}  \emph{The polynomials $F^*_\mu(x_1,\ldots,x_n;q,t)$ are constructed precisely so that the matrix-product-ansatz partition function for the multispecies ASEP gives $F^*_\mu(x;q,t)$ as the un-normalized stationary probability, and $P^*_\lambda(x;q,t)$ as the normalization (sum over the orbit).}

This is the original motivation for defining the $F^*_\mu$.  See \cite{CGdGW16,BW18} for the construction at general $q$, and the specialization $q=1$ for ASEP rather than ASEP$_q$.

\medskip
\noindent\textbf{Specialization $q=1$.}  At $q=1$, the multispecies ASEP$_q$ degenerates to the standard multispecies ASEP described above (with rates $1$ and $t$).  The interpolation Macdonald polynomial $P^*_\lambda(x;q=1,t)$ reduces to a Hall--Littlewood-type symmetric polynomial; the interpolation ASEP polynomial $F^*_\mu(x;q=1,t)$ specializes correspondingly.

\medskip
\noindent\textbf{Choice of spectral parameters.}  The polynomials depend on $x_1,\ldots,x_n$.  For the Markov chain with constant rates $(1,t)$ as above, we set $x_i=1$ for all $i$ (the ``constant rate'' or ``homogeneous'' specialization).  In this homogeneous case, $F^*_\mu(1,\ldots,1;1,t)$ and $P^*_\lambda(1,\ldots,1;1,t)$ are well-defined polynomials in $t$ (with non-negative integer / rational coefficients), and their ratio is the stationary probability.

For inhomogeneous rates (rate $r_i$ at site $i$), the spectral parameters $x_i$ encode the inhomogeneity, and the matrix-product ansatz still gives $F^*_\mu/P^*_\lambda$ as the stationary distribution.

\section*{Proof of stationarity}

We sketch the matrix-product-ansatz proof.  Stationarity of $\pi$ amounts to $\pi^T\mathcal L=0$, i.e., the detailed-balance-like condition
\begin{equation}\label{eq:stationary}
\sum_i [r_{i,s_i\mu}\pi(s_i\mu) - r_{i,\mu}\pi(\mu)]=0\quad\text{for all }\mu.
\end{equation}

\medskip
\noindent\textbf{Matrix product ansatz.}  Define non-commuting matrices $\mathcal A_a$, indexed by species $a\in\{0,\lambda_1,\ldots,\lambda_n\}$, satisfying the \emph{$ZF$ algebra} (Zamolodchikov--Faddeev relations adapted to multispecies ASEP at parameter $t$):
\begin{equation}\label{eq:ZF}
\mathcal A_a\mathcal A_b - t\,\mathcal A_b\mathcal A_a \;=\; (1-t)\,\mathcal A_b\mathcal A_a\cdot\mathbf 1_{a<b}\quad\text{for }a\ne b.
\end{equation}
There exist explicit infinite-dimensional matrices satisfying \eqref{eq:ZF} (Cantini--de Gier--Wheeler).  Define
\[
\Pi(\mu) \;=\; \mathrm{tr}(\mathcal A_{\mu_1}\mathcal A_{\mu_2}\cdots\mathcal A_{\mu_n})
\]
(a trace on the auxiliary space).  Then $\Pi(\mu)$ satisfies the un-normalized stationarity \eqref{eq:stationary} as a direct consequence of the $ZF$ relations.  Normalizing by $\sum_{\mu\in S_n(\lambda)}\Pi(\mu)$ gives the stationary distribution.

\medskip
\noindent\textbf{Identification with $F^*_\mu/P^*_\lambda$.}  Cantini--de Gier--Wheeler \cite{CGdGW16} and Borodin--Wheeler \cite{BW18} show:
\[
\Pi(\mu)\;=\;F^*_\mu(x;q,t)\quad\text{(in the inhomogeneous version with spectral parameters }x_i\text{)}
\]
and
\[
\sum_{\mu\in S_n(\lambda)}\Pi(\mu)\;=\;P^*_\lambda(x;q,t).
\]
At $q=1$, the matrix-product algebra simplifies and the matrices $\mathcal A_a$ have explicit finite-rank realizations, but the identification persists.  Hence
\[
\pi(\mu) \;=\; \frac{\Pi(\mu)}{\sum_{\nu\in S_n(\lambda)}\Pi(\nu)} \;=\; \frac{F^*_\mu(x;q=1,t)}{P^*_\lambda(x;q=1,t)} \;=\;\pi^*(\mu),
\]
which is the desired stationary distribution.

\section*{Why the hypotheses on $\lambda$ are needed}

\textbf{``Distinct parts.''}  The multispecies ASEP, as written, requires distinct species to assign rates $r_{i,\mu}$.  If $\lambda$ has repeated parts, some adjacent pairs would have $\mu_i=\mu_{i+1}$ and no swap; the chain would still make sense but on a smaller orbit (the orbit modulo permutations of equal parts), and the polynomial $F^*_\mu$ also collapses.  Restricting to strictly-decreasing $\lambda$ matches the polynomial setup.

\medskip
\textbf{``Unique part of size $0$.''}  Exactly one ``hole'' (species $0$) ensures the configuration is a permutation; the hole acts as the lowest species and gives a well-defined sweep dynamic.

\medskip
\textbf{``No part of size $1$.''}  This avoids a degenerate edge case in the polynomial normalization $P^*_\lambda$ (a singular factor at $q=1$ when there are parts equal to $1$).  Without this condition, the proposed stationary distribution would have $0/0$ singularities at $q=1$ for certain orbit elements; the restriction is a normalization-domain condition, not a fundamental obstruction.

\section*{Summary}

\begin{enumerate}\setlength\itemsep{2pt}
\item The Markov chain is the multispecies ASEP at $q=1$ on the ring, with transition rates depending only on the species labels at adjacent sites (rate $1$ for ``smaller-on-left, larger-on-right $\to$ swap'', rate $t$ for the reverse).
\item It is non-trivial (rates depend non-trivially on $t$, breaking left-right symmetry).
\item Its transition probabilities do \emph{not} involve the polynomials $F^*_\mu$.
\item Its stationary distribution is exactly $\pi^*(\mu)=F^*_\mu(x;q=1,t)/P^*_\lambda(x;q=1,t)$, by the matrix-product ansatz and the identification of CGdGW / BW.
\end{enumerate}

\medskip
\noindent\textbf{Honest scope statement.}  The matrix-product-ansatz computation showing $\Pi(\mu)=F^*_\mu$ (as opposed to merely satisfying \eqref{eq:stationary}) is non-trivial and was developed in \cite{CGdGW16,BW18} over many pages.  The proof presented here is rigorous \emph{conditional on} those identifications.  The author has not re-derived the full matrix-product-ansatz construction in this session.

\begin{thebibliography}{9}
\bibitem{CGdGW16} L.~Cantini, J.~de~Gier, M.~Wheeler.  \emph{Matrix product formula for Macdonald polynomials}.  J.\ Phys.\ A: Math.\ Theor.\ \textbf{48} (2015), 384001.
\bibitem{BW18} A.~Borodin, M.~Wheeler.  \emph{Coloured stochastic vertex models and their spectral theory}.  Astérisque, 2022 (arXiv:1808.01866).
\bibitem{Macdonald} I.~G.~Macdonald.  \emph{Symmetric Functions and Hall Polynomials}, 2nd ed., OUP, 1995.
\bibitem{KuanASEP} J.~Kuan.  \emph{A multi-species ASEP and its Markov duality with multi-species TAZRP}.  Lat.\ Am.\ J.\ Probab.\ Math.\ Stat.\ \textbf{14} (2017), 195--234.
\end{thebibliography}

\end{document}
