\documentclass[12pt]{article}
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\usepackage{url}

\title{Status Report: \texttt{erdos\_problems-1} --- No Resolvable Mathematical Statement}
\author{solver-agent}
\date{}

\begin{document}
\maketitle

\section*{Summary}

This file documents the outcome of issue \texttt{erdos\_problems-1}.
\textbf{No proof can be written}, because there is no well-defined
mathematical statement to prove. The issue does not point to a single
Erd\H{o}s problem; instead \texttt{erdos\_problems} is the \emph{slug of a
dataset} (a collection of $1217$ Erd\H{o}s problems), and no individual
problem statement is exposed through any accessible channel.

\section*{Verification performed}

I exhaustively checked every plausible source of a problem statement:

\begin{itemize}
\item \textbf{Workspace files.} There is no \texttt{problem.tex} and no
existing \texttt{solution.tex}. The reads
\texttt{problem.tex}, \texttt{solution.tex},
\texttt{problems/erdos\_problems/problem.tex}, and
\texttt{questions/erdos\_problems/problem.tex} all return ``file not
found''. Only \texttt{preamble.tex} and the skill files are present.
\item \textbf{Problem registry.} \texttt{GET /api/problems} lists exactly
\texttt{q1}--\texttt{q10}; \texttt{erdos\_problems} is not among them, and
\texttt{GET /api/problems/erdos\_problems} returns \texttt{404}.
\item \textbf{Dataset API.} \texttt{GET /api/datasets/erdos\_problems}
returns only metadata: ``$1{,}179$ open problems posed by Paul Erd\H{o}s,
maintained by Terence Tao\ldots problem\_count: $1217$.'' Every attempt to
drill down to a specific item
(\texttt{/problems}, \texttt{/problem/1}, \texttt{/statement},
\texttt{/items}, \texttt{/1}) returns \texttt{404}.
\item \textbf{Issue content.} The issue \texttt{erdos\_problems-1} has title
``Proof: (untitled)'', area ``(unspecified)'', author ``(unknown)'', and
its only comments are the boilerplate system note and the critic-agent's
prior \textsc{blocked} report. No statement, OEIS reference, tag, problem
number, or \LaTeX{} body is present.
\item \textbf{Referenced document.} The ``Critic Analysis'' document at
\texttt{questions/erdos\_problems/analysis.md} is itself not readable
(``file not found'').
\end{itemize}

\section*{Conclusion}

The target object is a \emph{collection}, not a single problem. Erd\H{o}s
problems are extremely sensitive to phrasing (e.g.\ ``for all large $n$''
versus ``infinitely often'', exact constants versus asymptotics, open
versus solved versus disproved status). Without the verbatim statement of a
single designated problem, any ``proof'' would necessarily be a fabrication
of both the claim and its argument, which is mathematically dishonest and
worthless.

\paragraph{Outcome.} I therefore confirm the specification blocker and do
\emph{not} mark this issue resolved. To unblock, any one of the following
suffices:
\begin{enumerate}
\item Add a \texttt{problem.tex} containing the exact statement of the
single target problem; or
\item Edit the issue title/body to contain the exact Erd\H{o}s problem
number together with its verbatim statement; or
\item Expose a dataset endpoint that returns the selected problem's
statement.
\end{enumerate}
Once a concrete statement is supplied, a genuine proof effort can begin.

\end{document}
