\documentclass{amsart}
\usepackage{amsmath,amssymb,amsthm,hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}

\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

\begin{document}

\title{On the computational complexity of maximum--likelihood mixture modeling}
\maketitle

\begin{abstract}
We give rigorous complexity results for the maximum--likelihood mixture--modeling
problem. We prove that, already for the most benign continuous family
$\mathcal F$ --- isotropic Gaussians with a fixed, prescribed covariance --- the
natural decision version of the problem is \textbf{NP--hard}, via a
many--one reduction from Euclidean $k$--means whose numerical parameters are of
polynomial bit size. As a consequence the optimization problem admits no
polynomial--time algorithm computing the optimal log--likelihood unless
$\mathrm{P}=\mathrm{NP}$, and the corresponding additive approximation problem is
NP--hard. On the positive side we show that for $k=1$ the problem is solvable in
closed form in polynomial time, and for constant $(d,k)$ it is polynomial; we thus
delineate precisely which features of the input are responsible for the hardness.
\end{abstract}

\section{Formalization}

\subsection{The mixture-modeling problem}
Let $\mathcal F=\{f(\cdot\mid\theta):\theta\in\Theta\}$ be a parametric family of
probability density functions on $\R^{d}$. Given a finite multiset
$X=\{x_{1},\dots,x_{n}\}\subset\R^{d}$ and an integer $k\ge1$, a
\emph{$k$-mixture} is a pair $(\theta,\pi)$ with parameters
$\theta=(\theta_{1},\dots,\theta_{k})\in\Theta^{k}$ and mixing weights
$\pi=(\pi_{1},\dots,\pi_{k})$ satisfying $\pi_{j}\ge0$ and
$\sum_{j=1}^{k}\pi_{j}=1$. Its log-likelihood is
\begin{equation}\label{eq:loglik}
L(\theta,\pi)\;=\;\sum_{i=1}^{n}\log\!\Big(\sum_{j=1}^{k}\pi_{j}\,f(x_{i}\mid\theta_{j})\Big),
\end{equation}
and we write $L^{\ast}(X,k,\mathcal F)=\sup_{(\theta,\pi)}L(\theta,\pi)$.

\subsection{The family used for the hardness result}
Fix a scale parameter $\sigma>0$. The \emph{fixed--covariance isotropic Gaussian
family} $\mathcal G_{\sigma}$ in dimension $d$ is parametrized by the mean
$\theta=m\in\R^{d}$, with density
\begin{equation}\label{eq:gauss}
f(x\mid m)\;=\;c_{\sigma}\,\exp\!\Big(-\tfrac{1}{2\sigma^{2}}\norm{x-m}^{2}\Big),
\qquad c_{\sigma}:=(2\pi\sigma^{2})^{-d/2},
\end{equation}
where $\norm{\cdot}$ is the Euclidean norm. Only the means $m_{1},\dots,m_{k}$ and
the weights $\pi$ are free; $\sigma$ is part of the input and is held fixed during
the optimization. Fixing $\sigma$ is both natural and necessary: if the covariance
scale is also free, the likelihood \eqref{eq:loglik} is unbounded above for
$k\ge2$ (Remark~\ref{rem:degen}), so the supremum problem would be ill-posed.

\begin{definition}[Mixture--modeling decision problem]\label{def:mmdec}
\textsc{GMM-MLE} is the following decision problem. \emph{Input:} a finite set
$X\subset\Z^{d}$ of $n$ points with integer coordinates, an integer $k\ge2$, a
rational scale $\sigma^{2}>0$, and a rational threshold $\tau$, all encoded in
binary. \emph{Question:} is $L^{\ast}(X,k,\mathcal G_{\sigma})\ge\tau$?
Because the additive constant $n\log c_{\sigma}$ is transcendental, we phrase the
threshold equivalently as a rational comparison on the \emph{normalized}
log-likelihood $L^{\ast}-n\log c_{\sigma}$; this removes any issue of representing
$\log$ and changes nothing below.
\end{definition}

\subsection{The combinatorial problem we reduce from}
\begin{definition}[Euclidean $k$-means]\label{def:kmeans}
For $X\subset\R^{d}$ and centers $M=(m_{1},\dots,m_{k})\in(\R^{d})^{k}$ put
\[
\mathrm{KM}(X,M)=\sum_{i=1}^{n}\min_{1\le j\le k}\norm{x_{i}-m_{j}}^{2},
\qquad
\mathrm{OPT}(X,k)=\inf_{M\in(\R^{d})^{k}}\mathrm{KM}(X,M).
\]
The decision problem \textsc{$k$-Means} is: given $X\subset\Z^{d}$, an integer
$k\ge2$, and a rational $T$, decide whether $\mathrm{OPT}(X,k)\le T$. The
\emph{integer-threshold} variant \textsc{$k$-Means$_{\Z}$} is the same problem
restricted to instances in which every achievable partition cost (in the sense of
Lemma~\ref{lem:centroid}) is an integer and $T\in\Z$.
\end{definition}

\begin{theorem}[Hardness, after integral scaling]\label{thm:kmnp}
\textsc{$k$-Means} is NP-hard, already for $k=2$ and unrestricted dimension $d$
\cite{ADHP}. Moreover \textsc{$k$-Means$_{\Z}$} is NP-hard for $k=2$ and
unrestricted $d$.
\end{theorem}

\begin{proof}
The first statement is \cite{ADHP}. For the second, we reduce \textsc{$k$-Means} to
\textsc{$k$-Means$_{\Z}$}. Let $(X,k,T)$ be a \textsc{$k$-Means} instance with
$X\subset\Z^{d}$, $|X|=n$, and $T=a/c$ in lowest terms ($a\in\Z$, $c\in\Z_{>0}$).
Put
\[
N:=\mathrm{lcm}(1,2,\dots,n).
\]
By the prime-number theorem $\log N=\sum_{p\le n}\lfloor\log_{p}n\rfloor\log p
=O(n)$, so $N$ has $O(n)$ bits and is computable in polynomial time. Define the
scaled data $X':=\{N\,x_{1},\dots,N\,x_{n}\}\subset\Z^{d}$ (still polynomial input
size, each coordinate gaining $O(n)$ bits) and the integer threshold
\[
T'':=\big\lfloor N^{2}T\big\rfloor\in\Z .
\]
Scaling all coordinates by $N$ multiplies every squared distance, hence every
$k$-means cost, by $N^{2}$:
\(
\mathrm{OPT}(X',k)=N^{2}\,\mathrm{OPT}(X,k).
\)
By Lemma~\ref{lem:sep}, every achievable cost of $X$ is $\sum_{l}A_{l}/|P_{l}|$
with $A_{l}\in\Z_{\ge0}$ and $|P_{l}|\le n$; hence the corresponding cost of $X'$ is
$\sum_{l}N^{2}A_{l}/|P_{l}|$, and since $|P_{l}|\mid N$ each summand
$N^{2}A_{l}/|P_{l}|=A_{l}\,N\,(N/|P_{l}|)$ is an integer. Thus all achievable costs
of $X'$ are integers, so $(X',k,T'')$ is a valid \textsc{$k$-Means$_{\Z}$}
instance. Finally
\[
\mathrm{OPT}(X,k)\le T
\iff N^{2}\,\mathrm{OPT}(X,k)\le N^{2}T
\iff \mathrm{OPT}(X',k)\le \lfloor N^{2}T\rfloor=T'',
\]
the last equivalence because $\mathrm{OPT}(X',k)$ is an integer (so
$\mathrm{OPT}(X',k)\le N^{2}T$ iff $\mathrm{OPT}(X',k)\le\lfloor N^{2}T\rfloor$).
Hence the reduction is correct and polynomial, and \textsc{$k$-Means$_{\Z}$} is
NP-hard.
\end{proof}

\section{Structure of the $k$-means optimum}

\begin{lemma}[Centroid form]\label{lem:centroid}
Let $X\subset\R^{d}$. For a partition $\{P_{1},\dots,P_{k}\}$ of $\{1,\dots,n\}$
set $\mu_{l}=\frac{1}{|P_{l}|}\sum_{i\in P_{l}}x_{i}$ for nonempty $P_{l}$. Then
for any single center $m$ and nonempty $P$,
\begin{equation}\label{eq:par}
\sum_{i\in P}\norm{x_{i}-m}^{2}=\sum_{i\in P}\norm{x_{i}-\mu}^{2}+|P|\,\norm{m-\mu}^{2},
\qquad \mu=\tfrac{1}{|P|}\textstyle\sum_{i\in P}x_{i},
\end{equation}
so the cost of any cluster is minimized uniquely at its centroid, and
\[
\mathrm{OPT}(X,k)=\min_{\{P_{l}\}}\ \sum_{l=1}^{k}\sum_{i\in P_{l}}\norm{x_{i}-\mu_{l}}^{2},
\]
the minimum over all partitions of $\{1,\dots,n\}$ into at most $k$ nonempty parts,
and it is attained.
\end{lemma}

\begin{proof}
Expanding the left side of \eqref{eq:par},
$\sum_{i\in P}\norm{x_{i}-m}^{2}=\sum_{i\in P}\norm{x_{i}-\mu}^{2}
+2\langle\mu-m,\sum_{i\in P}(x_{i}-\mu)\rangle+|P|\norm{m-\mu}^{2}$, and the middle
term vanishes since $\sum_{i\in P}(x_{i}-\mu)=0$. This gives \eqref{eq:par} and the
unique minimizer $m=\mu$. For the global statement, $\mathrm{KM}(X,M)$ equals, for
the Voronoi partition induced by $M$, the sum $\sum_{l}\sum_{i\in P_{l}}
\norm{x_{i}-m_{l}}^{2}$, and conversely every partition is a feasible assignment;
minimizing over centers (centroids) then over the finitely many partitions yields
the formula, with the infimum attained.
\end{proof}

\begin{lemma}[Separation of achievable costs]\label{lem:sep}
Let $X\subset\Z^{d}$ with $|X|=n$. Every partition cost
$\sum_{l}\sum_{i\in P_{l}}\norm{x_{i}-\mu_{l}}^{2}$ is a rational number of the form
$\sum_{l}A_{l}/|P_{l}|$ with $A_{l}\in\Z_{\ge0}$, hence its reduced denominator is
$\le n$. Consequently any two \emph{distinct} achievable partition costs $u\ne v$
satisfy $|u-v|\ge 1/n^{2}$.
\end{lemma}

\begin{proof}
For a nonempty part $P$ with $m:=|P|$, by \eqref{eq:par} with the center equal to
$0$ rearranged, $\sum_{i\in P}\norm{x_{i}-\mu}^{2}=\sum_{i\in P}\norm{x_{i}}^{2}
-m\norm{\mu}^{2}$. Using $m\mu=\sum_{i\in P}x_{i}$,
\[
\sum_{i\in P}\norm{x_{i}-\mu}^{2}
=\frac{1}{m}\Big(m\sum_{i\in P}\norm{x_{i}}^{2}-\Big\Vert\sum_{i\in P}x_{i}\Big\Vert^{2}\Big)
=\frac{A_{P}}{m},
\]
where $A_{P}=m\sum_{i\in P}\norm{x_{i}}^{2}-\norm{\sum_{i\in P}x_{i}}^{2}$ is an
integer (the $x_{i}$ are integer vectors) and nonnegative (it is $m$ times the
left side, which is nonnegative). Summing over parts, each partition cost is
$\sum_{l}A_{l}/|P_{l}|$, a rational with denominator dividing
$\mathrm{lcm}(1,\dots,n)$; after reduction its denominator is at most $n$. Writing
$u=p/q$, $v=r/s$ in lowest terms with $1\le q,s\le n$, if $u\ne v$ then
$u-v=(ps-rq)/(qs)$ has a nonzero integer numerator, so $|u-v|\ge1/(qs)\ge1/n^{2}$.
\end{proof}

\begin{corollary}[Clean integer dichotomy]\label{cor:dich}
Let $(X,k,T)$ be a \textsc{$k$-Means$_{\Z}$} instance: $X\subset\Z^{d}$, every
achievable partition cost an integer, $T\in\Z$. Then $\mathrm{OPT}(X,k)\in\Z$, and
exactly one of the following holds:
\[
\textup{(a) } \mathrm{OPT}(X,k)\le T,
\qquad\text{or}\qquad
\textup{(b) } \mathrm{OPT}(X,k)\ge T+1 .
\]
\end{corollary}

\begin{proof}
By Lemma~\ref{lem:centroid}, $\mathrm{OPT}(X,k)$ is an achievable partition cost,
hence an integer by hypothesis. If $\mathrm{OPT}(X,k)>T$ with both integers, then
$\mathrm{OPT}(X,k)\ge T+1$, giving (b); otherwise (a) holds. The cases are mutually
exclusive.
\end{proof}

\section{The likelihood--cost sandwich}

\begin{lemma}[Sandwich]\label{lem:sandwich}
Fix $\sigma>0$, $k\ge2$, $X\subset\R^{d}$ with $|X|=n$, and
$c_{\sigma}=(2\pi\sigma^{2})^{-d/2}$. With $\mathcal G_{\sigma}$ as in
\eqref{eq:gauss},
\begin{equation}\label{eq:sandwich}
n\log c_{\sigma}-\frac{\mathrm{OPT}(X,k)}{2\sigma^{2}}-n\log k
\;\le\;
L^{\ast}(X,k,\mathcal G_{\sigma})
\;\le\;
n\log c_{\sigma}-\frac{\mathrm{OPT}(X,k)}{2\sigma^{2}}.
\end{equation}
\end{lemma}

\begin{proof}
\emph{Upper bound.} Let $(\theta,\pi)=((m_{1},\dots,m_{k}),\pi)$ be any
$k$-mixture. For each $i$, since $e^{-t}$ is decreasing in $t\ge0$ and
$\sum_{j}\pi_{j}=1$ with $\pi_{j}\ge0$,
\[
\sum_{j=1}^{k}\pi_{j}\,c_{\sigma}e^{-\norm{x_{i}-m_{j}}^{2}/(2\sigma^{2})}
\;\le\;c_{\sigma}\Big(\max_{j}e^{-\norm{x_{i}-m_{j}}^{2}/(2\sigma^{2})}\Big)\sum_{j}\pi_{j}
\;=\;c_{\sigma}\,e^{-\min_{j}\norm{x_{i}-m_{j}}^{2}/(2\sigma^{2})}.
\]
Taking logarithms and summing over $i$,
\[
L(\theta,\pi)\le\sum_{i=1}^{n}\Big(\log c_{\sigma}-\frac{\min_{j}\norm{x_{i}-m_{j}}^{2}}{2\sigma^{2}}\Big)
=n\log c_{\sigma}-\frac{\mathrm{KM}(X,M)}{2\sigma^{2}}
\le n\log c_{\sigma}-\frac{\mathrm{OPT}(X,k)}{2\sigma^{2}},
\]
using $\mathrm{KM}(X,M)\ge\mathrm{OPT}(X,k)$. Taking the supremum over mixtures
gives the right inequality of \eqref{eq:sandwich}.

\emph{Lower bound.} By Lemma~\ref{lem:centroid} there are centers $M^{\star}$ with
$\mathrm{KM}(X,M^{\star})=\mathrm{OPT}(X,k)$. Use the mixture with means
$M^{\star}$ and uniform weights $\pi_{j}=1/k$. Keeping only the nearest center
$j^{\star}(i)\in\arg\min_{j}\norm{x_{i}-m^{\star}_{j}}^{2}$,
\[
\sum_{j=1}^{k}\frac1k c_{\sigma}e^{-\norm{x_{i}-m^{\star}_{j}}^{2}/(2\sigma^{2})}
\;\ge\;\frac{c_{\sigma}}{k}\,e^{-\min_{j}\norm{x_{i}-m^{\star}_{j}}^{2}/(2\sigma^{2})}.
\]
Taking logarithms and summing,
\[
L^{\ast}\ge L(M^{\star},\pi)\ge n\log c_{\sigma}-n\log k-\frac{\mathrm{OPT}(X,k)}{2\sigma^{2}},
\]
the left inequality of \eqref{eq:sandwich}.
\end{proof}

\section{NP-hardness}

\begin{theorem}[Hardness of mixture modeling]\label{thm:main}
\textsc{GMM-MLE} (Definition~\ref{def:mmdec}) is NP-hard. Hence, unless
$\mathrm{P}=\mathrm{NP}$, no polynomial-time algorithm computes
$L^{\ast}(X,k,\mathcal G_{\sigma})$ (equivalently, the normalized optimum
$L^{\ast}-n\log c_{\sigma}$) exactly for inputs with $k\ge2$.
\end{theorem}

\begin{proof}
We reduce \textsc{$k$-Means$_{\Z}$} (NP-hard by Theorem~\ref{thm:kmnp}) to
\textsc{GMM-MLE}.

\emph{Reduction.} Let $(X,k,T)$ be a \textsc{$k$-Means$_{\Z}$} instance:
$X\subset\Z^{d}$, $|X|=n$, $k\ge2$, $T\in\Z$, with all achievable partition costs
integral. By Corollary~\ref{cor:dich} exactly one of $\mathrm{OPT}(X,k)\le T$ and
$\mathrm{OPT}(X,k)\ge T+1$ holds; the gap is $\delta:=1$.

\emph{Choice of scale and threshold.} The ideal scale is
$\sigma^{2}=1/(4n\log k)$. To keep the encoding rational, replace $\log k$ by a
rational \emph{upper} approximation $L_{k}\in\Q$ with
\begin{equation}\label{eq:Lk}
\log k\;\le\;L_{k}\;\le\;2\log k,
\end{equation}
obtained by rounding $\log k$ up to $b:=\lceil 2\log_{2}n\rceil$ binary places
(possible since $\log k>0$ for $k\ge2$). Set
\[
\sigma^{2}:=\frac{1}{4\,n\,L_{k}},
\qquad
\widehat\tau:=-\big(2nL_{k}\big)\Big(T+\tfrac12\Big),
\]
so that, with $1/(2\sigma^{2})=2nL_{k}$,
\begin{equation}\label{eq:tauform}
\widehat\tau=-\frac{T+\tfrac12\delta}{2\sigma^{2}}\qquad(\delta=1).
\end{equation}
The output of the reduction is the \textsc{GMM-MLE} instance
$(X,k,\sigma^{2},\widehat\tau)$, asking whether
$L^{\ast}-n\log c_{\sigma}\ge\widehat\tau$ (the normalized comparison of
Definition~\ref{def:mmdec}).

\emph{Polynomial size.} $\sigma^{2}=1/(4nL_{k})$ and $\widehat\tau$ are rationals
of bit length $O(\log n+\mathrm{size}(T))$, computable in polynomial time. The one
arithmetic fact used below is, from \eqref{eq:Lk} with $\delta=1$,
\begin{equation}\label{eq:key}
\frac{\delta}{2\sigma^{2}}=\frac{1}{2}\cdot 4nL_{k}=2nL_{k}\ge 2n\log k .
\end{equation}

\emph{Correctness.}

\noindent(1) Suppose $\mathrm{OPT}(X,k)\le T$. By the lower bound in
\eqref{eq:sandwich},
\[
L^{\ast}-n\log c_{\sigma}
\;\ge\;-\frac{\mathrm{OPT}(X,k)}{2\sigma^{2}}-n\log k
\;\ge\;-\frac{T}{2\sigma^{2}}-n\log k .
\]
This is $\ge\widehat\tau=-\dfrac{T+\delta/2}{2\sigma^{2}}$ iff
$\dfrac{\delta/2}{2\sigma^{2}}\ge n\log k$, i.e.\ iff
$\dfrac{\delta}{2\sigma^{2}}\ge 2n\log k$, which is exactly \eqref{eq:key}. Hence
$L^{\ast}-n\log c_{\sigma}\ge\widehat\tau$.

\noindent(2) Suppose $\mathrm{OPT}(X,k)\ge T+\delta$. By the upper bound in
\eqref{eq:sandwich},
\[
L^{\ast}-n\log c_{\sigma}\le-\frac{\mathrm{OPT}(X,k)}{2\sigma^{2}}
\le-\frac{T+\delta}{2\sigma^{2}}
<-\frac{T+\delta/2}{2\sigma^{2}}=\widehat\tau,
\]
the strict inequality because $\delta>\delta/2$ and $\sigma^{2}>0$.

Cases (1)--(2) give
\(
\mathrm{OPT}(X,k)\le T \iff L^{\ast}-n\log c_{\sigma}\ge\widehat\tau,
\)
so the map $(X,k,T)\mapsto(X,k,\sigma^{2},\widehat\tau)$ is a correct
polynomial-time many-one reduction; thus \textsc{GMM-MLE} is NP-hard. If $L^{\ast}$
(equivalently $L^{\ast}-n\log c_{\sigma}$) were computable in polynomial time,
comparison with $\widehat\tau$ would decide \textsc{GMM-MLE}, hence
\textsc{$k$-Means$_{\Z}$}, forcing $\mathrm{P}=\mathrm{NP}$.
\end{proof}

\begin{remark}
The rational approximation $L_{k}$ of $\log k$ serves only to keep the encoded
scale $\sigma^{2}$ and threshold $\widehat\tau$ rational. The choice
$L_{k}\ge\log k$ is what makes the penalty identity \eqref{eq:key} hold; the
upper-bound case (2) uses no property of $L_{k}$. Thus the reduction is exact and
polynomial.
\end{remark}

\section{Hardness of approximation}

We measure approximation additively on the log-likelihood scale, the intrinsic
scale of \eqref{eq:loglik}.

\begin{corollary}[Additive inapproximability]\label{cor:approx}
Unless $\mathrm{P}=\mathrm{NP}$, no polynomial-time algorithm, given
$(X,k,\sigma^{2})$ with $X\subset\Z^{d}$, $k\ge2$, outputs a value $\widetilde L$
with $\big|\widetilde L-\big(L^{\ast}(X,k,\mathcal G_{\sigma})-n\log c_{\sigma}\big)\big|
<\tfrac12 n\log k$; that is, the normalized optimum cannot be approximated within
additive error $\tfrac12 n\log k$ in polynomial time.
\end{corollary}

\begin{proof}
Run the reduction of Theorem~\ref{thm:main} from a \textsc{$k$-Means$_{\Z}$}
instance $(X,k,T)$ (so $\delta=1$), but halve the scale: set
$\sigma^{2}:=1/(8nL_{k})$ with $L_{k}\ge\log k$ as in \eqref{eq:Lk}, so
$\tfrac{1}{2\sigma^{2}}=4nL_{k}$. Write $V:=L^{\ast}-n\log c_{\sigma}$ for the
normalized optimum. By Corollary~\ref{cor:dich} we are in one of two cases.

If $\mathrm{OPT}(X,k)\le T$: by \eqref{eq:sandwich},
$V\ge -\tfrac{T}{2\sigma^{2}}-n\log k=:U$.

If $\mathrm{OPT}(X,k)\ge T+1$: by \eqref{eq:sandwich},
$V\le -\tfrac{T+1}{2\sigma^{2}}=U-\big(\tfrac{1}{2\sigma^{2}}-n\log k\big)$. Since
$\tfrac{1}{2\sigma^{2}}=4nL_{k}\ge 4n\log k$, we get
$\tfrac{1}{2\sigma^{2}}-n\log k\ge 3n\log k$, hence $V\le U-3n\log k$.

Thus the two cases are separated by an additive gap of at least $3n\log k$ in $V$,
around the efficiently computable pivot $U-\tfrac32 n\log k$ (note $U$ is rational
because the transcendental term $n\log c_{\sigma}$ has been removed). An algorithm
with additive error $<\tfrac12 n\log k<\tfrac32 n\log k$ on $V$ would, on comparing
$\widetilde L$ to $U-\tfrac32 n\log k$, correctly distinguish the cases and hence
decide \textsc{$k$-Means$_{\Z}$}, contradicting $\mathrm{P}\ne\mathrm{NP}$.
\end{proof}

\begin{remark}[Multiplicative form]
The affine reparametrization $\Phi:=n\log c_{\sigma}-L^{\ast}
=\mathrm{OPT}(X,k)/(2\sigma^{2})+E$ with $0\le E\le n\log k$ (from
\eqref{eq:sandwich}) is a nonnegative quantity whose exact computation is NP-hard
on the separated instances of Lemma~\ref{lem:sep}. Since on those instances
$\mathrm{OPT}/(2\sigma^{2})$ values for the two cases differ by a factor bounded
away from $1$ while $E$ is comparatively negligible, computing $\Phi$ within a
suitable constant multiplicative factor would recover the boolean
$\mathrm{OPT}\le T$ and is therefore NP-hard. Hence there is a constant
approximation threshold below which the (normalized) problem is intractable.
\end{remark}

\section{The tractable boundary}

\begin{proposition}[$k=1$ is in P]\label{prop:k1}
Let $k=1$.
\begin{enumerate}
\item[(i)] For $\mathcal G_{\sigma}$ the unique optimal mean is the sample mean
$\bar x=\frac1n\sum_{i}x_{i}$, the only weight is $\pi_{1}=1$, and
\[
L^{\ast}(X,1,\mathcal G_{\sigma})=n\log c_{\sigma}-\frac{1}{2\sigma^{2}}\sum_{i=1}^{n}\norm{x_{i}-\bar x}^{2},
\]
computable in $O(nd)$ arithmetic operations.
\item[(ii)] For the full isotropic Gaussian (mean and $\sigma$ free), the MLE is
$m=\bar x$, $\sigma^{2}=\frac{1}{nd}\sum_{i}\norm{x_{i}-\bar x}^{2}$, provided the
data are not all equal, with optimal value computable in $O(nd)$ operations.
\end{enumerate}
\end{proposition}

\begin{proof}
(i) With $k=1$ we have $\pi_{1}=1$, so
$L(m)=n\log c_{\sigma}-\frac{1}{2\sigma^{2}}\sum_{i}\norm{x_{i}-m}^{2}$. By
\eqref{eq:par} with $P=\{1,\dots,n\}$, $\sum_{i}\norm{x_{i}-m}^{2}
=\sum_{i}\norm{x_{i}-\bar x}^{2}+n\norm{m-\bar x}^{2}$, minimized uniquely at
$m=\bar x$. Substituting gives the value, computed in one pass.

(ii) For fixed $\sigma$, the inner maximization over $m$ is at $m=\bar x$ by (i).
Let $S=\sum_{i}\norm{x_{i}-\bar x}^{2}>0$ and $t=\sigma^{2}$; maximize
$h(t)=-\tfrac{nd}{2}\log(2\pi t)-\tfrac{S}{2t}$. Then
$h'(t)=-\tfrac{nd}{2t}+\tfrac{S}{2t^{2}}$, vanishing uniquely at $t=S/(nd)$, where
$h''(t)=\tfrac{nd}{2t^{2}}-\tfrac{S}{t^{3}}=\tfrac{1}{t^{2}}(\tfrac{nd}{2}-\tfrac{S}{t})
=\tfrac{1}{t^{2}}(\tfrac{nd}{2}-nd)<0$, so it is the unique maximizer. All
quantities are computable in $O(nd)$ operations.
\end{proof}

\begin{proposition}[Constant $(d,k)$ is in P, up to any additive accuracy]\label{prop:dk}
For fixed $d$ and $k$, $\mathrm{OPT}(X,k)$ is computable exactly in time
$n^{O(dk)}$, and hence $L^{\ast}(X,k,\mathcal G_{\sigma})$ is determined to within
the additive slack $n\log k$ of \eqref{eq:sandwich} in time $n^{O(dk)}\cdot
\mathrm{poly}(\text{input})$.
\end{proposition}

\begin{proof}
By Lemma~\ref{lem:centroid}, $\mathrm{OPT}(X,k)$ equals the minimum of the
within-cluster sum-of-squares (with centroids as centers) over all partitions of
$X$ into at most $k$ clusters, and an optimal partition is the one induced by the
Voronoi diagram of the $k$ optimal centroids: each data point is assigned to its
nearest centroid. Hence it suffices to enumerate all partitions of $X$ that can
arise as the Voronoi partition of \emph{some} $k$-tuple of points in $\R^{d}$, and
to take the best.

We bound the number of such partitions. Two centroids $\mu_{l},\mu_{l'}$ separate
the data along the perpendicular-bisector hyperplane
$H_{ll'}=\{z:\langle z,\mu_{l}-\mu_{l'}\rangle=\tfrac12(\norm{\mu_{l}}^{2}-\norm{\mu_{l'}}^{2})\}$,
so the Voronoi partition of $X$ is determined by, for each of the $\binom{k}{2}$
ordered pairs, the side of $H_{ll'}$ on which each data point lies, i.e.\ by the
sign pattern of the $n$ data points with respect to $\binom{k}{2}$ hyperplanes in
$\R^{d}$. The map $(\mu_{1},\dots,\mu_{k})\mapsto$ (this sign pattern) factors
through the labeling $z\mapsto(\langle z,\mu_{l}-\mu_{l'}\rangle-\tfrac12(\norm{\mu_l}^2-\norm{\mu_{l'}}^2))_{l<l'}$,
which for fixed $z$ is an affine function of the parameter vector
$(\mu_{1},\dots,\mu_{k},\norm{\mu_1}^2,\dots,\norm{\mu_k}^2)\in\R^{(d+1)k}$. By the
classical bound on the number of sign patterns realized by $m$ affine functionals
in $\R^{p}$ (Warren's / the Sauer--Shelah bound for halfspaces of VC dimension
$p$), the number of distinct sign patterns --- hence the number of distinct
Voronoi partitions of $X$ by $k$ centers --- is at most
$\big(\tfrac{e\,n\binom{k}{2}}{(d+1)k}\big)^{(d+1)k}=n^{O(dk)}$. This is the
Inaba--Katoh--Imai bound \cite{IKI}.

Therefore: enumerate one representative partition for each realizable sign pattern
(e.g.\ by sweeping over the $n^{O(dk)}$ combinatorially distinct centroid
configurations, each a choice of which data points are co-decisive for the
$\binom{k}{2}$ bisectors), compute its centroid cost via Lemma~\ref{lem:centroid}
in $O(nd)$ time, and return the minimum; the optimal Voronoi partition is among
them, so the result equals $\mathrm{OPT}(X,k)$ exactly. The total time is
$n^{O(dk)}$. Substituting into \eqref{eq:sandwich} determines $L^{\ast}$ to within
the additive slack $n\log k$.
\end{proof}

\begin{remark}[Why $\sigma$ must be fixed]\label{rem:degen}
For $k\ge2$ with \emph{free} per-component covariances the supremum in
\eqref{eq:loglik} is $+\infty$: take component $1$ with mean $m_{1}=x_{1}$ and
covariance $s\,I$, $s\to0^{+}$, and a second component with a fixed positive
covariance whose density is bounded below by a positive constant on the data.
Then $\pi_{1}f(x_{1}\mid m_{1})=\pi_{1}(2\pi s)^{-d/2}\to+\infty$, so the
$i=1$ term of \eqref{eq:loglik} $\to+\infty$ while the others stay bounded below.
Thus a well-posed MLE for $k\ge2$ requires fixing the covariance scale (as in
$\mathcal G_{\sigma}$) or lower-bounding covariance eigenvalues; our hardness
applies to the most restrictive such regularization, and therefore is the
strongest possible in this respect.
\end{remark}

\section{Summary of the parameter regimes}

\begin{itemize}
\item \textbf{Number of components $k$.} Intractability appears already at $k=2$
(Theorem~\ref{thm:main}); $k=1$ is closed-form (Proposition~\ref{prop:k1}). The
component count is the primary driver of hardness.
\item \textbf{Dimension $d$.} The reduction needs $d$ unbounded, matching the
$k=2$ hardness of $k$-means. For constant $(d,k)$ the problem is polynomial
(Proposition~\ref{prop:dk}); the hardness is a high-dimensional phenomenon.
\item \textbf{Family $\mathcal F$.} Hardness holds for the most benign smooth
family --- isotropic Gaussians of a single fixed scale. Enlarging $\mathcal F$ to
free covariances only makes the problem harder or ill-posed
(Remark~\ref{rem:degen}); $k=1$ exponential-family fitting is tractable by
concavity of the single-component log-likelihood.
\item \textbf{Separation.} The sandwich \eqref{eq:sandwich} localizes the
difficulty to the regime where the optimal $k$-means partition is ambiguous, i.e.\
poorly separated components relative to $\sigma$. The slack $n\log k$ (a
soft-assignment entropy term) is dominated by the inter-cost separation
$\ge\sigma^{-2}n^{-2}$ on the hard instances, which is precisely what powers the
reduction; with strong separation the assignment becomes unambiguous and the
problem is tractable.
\item \textbf{Approximation threshold.} Computing $L^{\ast}$ within additive error
$\tfrac12 n\log k$ on the log-likelihood scale is NP-hard
(Corollary~\ref{cor:approx}); equivalently the normalized optimum admits no
polynomial-time approximation below a constant multiplicative factor unless
$\mathrm{P}=\mathrm{NP}$.
\end{itemize}

\begin{thebibliography}{9}
\bibitem{ADHP} D.~Aloise, A.~Deshpande, P.~Hansen, P.~Popat,
\emph{NP-hardness of Euclidean sum-of-squares clustering},
Machine Learning \textbf{75} (2009), 245--248.
\bibitem{IKI} M.~Inaba, N.~Katoh, H.~Imai,
\emph{Applications of weighted Voronoi diagrams and randomization to
variance-based $k$-clustering}, Proc.\ 10th Annual Symposium on Computational
Geometry (SoCG), 1994, 332--339.
\end{thebibliography}

\end{document}
